### 2009-06-16 paper.tex
last edits

We implemented the above algorithms for a simple model of pairwise RNA structural evolution;
in particular the algorithms for maximum likelihood (ML) alignment of three known RNA structures an a known phylogeny
and inference of the common ancestral structure.
We compared this ML algorithm to a variety of related, but simpler, techniques,
including ML alignment algorithms for simpler models that omitted various aspects of the full model
and also a heuristic (sum-over-pairs) posterior-decoding alignment algorithm for one of the simpler models.

Our results suggest that ML algorithms incorporating basepair structure are the most likely to estimate the alignment perfectly.
However, the heuristic posterior-decoding algorithm scores higher on finer-grained alignment accuracy metrics.
Since the posterior-decoding heuristic is substantially faster than exact phylogenetic ML,
this further motivates the use of sum-over-pairs (and approximate sum-over-pairs) heuristics
for paleogenetics and other alignment problems.

>>>>

We implemented the above algorithms for a simple model of pairwise RNA structural evolution;
in particular the algorithms for maximum likelihood (ML) alignment of three known RNA structures an a known phylogeny
and inference of the common ancestral structure.
We compared this ML algorithm to a variety of related, but simpler, techniques,
including ML alignment algorithms for simpler models that omitted various aspects of the full model
and also a posterior-decoding alignment algorithm for one of the simpler models.

In our tests, incorporation of basepair structure was the most important factor for alignment accuracy;
appropriate use of posterior-decoding was the next most important factor;
and fine details of the model were least important.
Since posterior-decoding heuristics can be substantially faster than exact phylogenetic ML,
this motivates the use of sum-over-pairs (and approximate sum-over-pairs) heuristics
for paleogenetics and other alignment problems.
